Who does big team science?

### 2025-01-14 manuscript_definition.Rmd
updating with new manuscript information

header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

@@

# things to import
var_list <- c("author_stats", "paper_stats", "author_paper_year",
              "career_time")

@@

outline:

-   what is team science / collaboration
-   research on how it's grown / changed over time?
-   Any research diversification?
    -   on who is publishing
    -   on where is represented
-   "big" teams - unanswered question ... people are now using term "big
    team" but it's not defined - can we define it and see if it's any
    different or just a product of growth/internet over time?

SAVS NEW VERSION!

Collaboration in scientific endeavors involves multiple researchers at
(potentially) multiple institutions to communicate and work together to
advance knowledge in their chosen field. Collaboration can manifest
uniquely in each project dependent on the skill sets, hypotheses, and
perspectives of collaborators. One key aspect of collaboration is the
flexibility allotted as it is shaped by the needs of the project and the
participants involved.

While collaboration is not new in science, interest in "team science" is
growing as individual researchers seek an interdisciplinary approach to
research or bring on more students to their project. Unlike general
collaboration, team science involves structured roles, coordinated
workflows, and shared resources to address challenges that would be
difficult for individual researchers or one small team to solve
independently. Team science often brings in some aspect of diversity but
likely is still very homogenous in several aspects.

The evolution of team science reflects broader shifts in research
practices, driven by two sources: 1) increasing globalization and
technology that allows for real-time interdisciplinary research, and 2)
expanding interest in reproducibility, replication, and generalizability
[i.e., the credibility movement, @maxwell2015; @nelson2018; @zwaan2018;
@vazire2022]. Technological advances have provided easier ways to
collaborate with people who are from other universities and countries
through document sharing platforms (e.g., Google, GitHub, and the Open
Science Framework), video chatting platforms (e.g., Zoom, Microsoft
Teams), and messaging and project management platforms (e.g., Slack,
Trello, when2meet, etc.).

The credibility movement has been the main catalyst for the increased
interest and broader shift of research practices. The movement
emphasizes reproducibility and transparency in science, encouraging
researchers to form new ways to increase the rigor in scientific
endeavors. As the movement has adapted throughout the last decade, the
idea to form larger, more diverse teams and to include participants from
varied backgrounds has become even more a goal to not only increase the
credibility but the generalizability and reliability of scientific
findings. This form of collaboration has been coined “Big Team Science.”

“Big Team Science” builds on team science by scaling efforts to include
larger, often globally diverse teams, which requires significant
coordination and infrastructure [@coles2022; @forscher2022a;
@stewart2017]. BTS projects and organizations organize extensive
collaborations, intentionally incorporating diverse populations and
perspectives into research. This large-scale approach enhances the
reliability and generalizability of findings by integrating varied
methodologies and viewpoints, leading to more robust and inclusive
scientific outcomes. BTS organizations often pool extensive networks of
researchers and resources, aiming to tackle grand scientific challenges
that would be difficult to address within smaller or less coordinated
collaborations. By having both collaborations that span across the globe
and subfields of research areas, age groups, and education levels should
help to drive science in the path of better materials, reliability,
generalizability, and more robust sample sizes (when necessary) in a
study [@auspurg2021; @nosek2014method; @lebel2018].

Generally, the credibility movement appears to have been driven by early
career researchers [i.e., those who are within five years of their first
appointment, @maizey2019]; however, there are no large meta-scientific
investigations on this specific topic to date. The newness of
large-scale research in many fields could be the culprit for the lack of
investigation into this area. For example, psychology has had an
increase in BTS publications like the Open Science Collaboration
[@opensciencecollaboration2015], Many Labs Collaborations
[@buttrick2020; @ebersole2016; @ebersole2020; @klein2018; @klein2022;
@mathur2020; @skorb2020] or the first papers from the Psychological
Science Accelerator [@bago2022; @buchanan2023; @dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@moshontz2018; @wang2021]. Generally, the researcher incentive for
replication and/or involvement in big-team projects was low for three
reasons. First, journals often prioritize "novel" or new results which
led to rejection of replication manuscripts and publication bias
[@franco2014; @hubbard1997; @nosek2012]. Second, the "failure" to
replicate was often placed on the replication team as "bad science"
rather than a careful consideration of publication biases and
(potential) questionable research practices [@klein2022; @maxwell2015].
Last, why should someone want to spend time and resources on an answer
we already "know" [@isager2021; @isager2021a]?

However, the success and interest in the large-scale reproducibility
projects [@opensciencecollaboration2015; @10.7554/eLife.71601], paired
with the meta-scientific publications focusing on researcher practices
and incentive structures [@silberzahn2018many; @john2012] led to a
change in journal guidelines and incentives for researchers interested
in participating in large-scale replication studies [@grahe2014;
@kidwell2016; @nosek2015; @mayo-wilson2021]. In some fields, the
replication movement demonstrated that large-scale teams were a
practical (and publishable) solution to answering research questions in
generalizable way. The support for Registered Reports, papers accepted
before the data has been collected [@nosek2014; @stewart2020], has
allowed researchers to invest in projects that they know should be
published when the project is complete. Further, the implementation of
the Transparency and Openness Guidelines [@nosek2015] and the
Contributor Role Taxonomy (CRediT) system [@allen2019] have pushed
journals and researchers to promote more open, inclusive publication
practices.

Beyond replication concerns, the credibility movement has mirrored calls
for diversification or de-WEIRDing (e.g., Western, Educated,
Industrialized, Rich, and Democratic) scientific research [@henrich2010;
@newson2021; @rad2018] by improving representation in research samples.
Like the large-scale studies in Physics [@aphilos2021; @castelnovo2018]
and Biology [@collins2003], the Social Sciences struggle to represent
the breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator [@moshontz2018], ManyBabies
(<https://manybabies.github.io/>), NutNet (<https://nutnet.org/>), and
DRAGNet (<https://dragnetglobal.weebly.com/>) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more [@bago2022;
@dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@vanbavel2022; @wang2021]. While these organizations and one-time groups
for BTS studies have provided an incredible wealth of data for the
scientific community, we do not yet know exactly *who* is involved with,
and benefits from, the BTS and credibility movement. Publications on BTS
generally explore challenges, lessons learned, and the need for BTS
[@forscher2022a; @coles2022].

There is no universally agreed-upon definition of BTS, raising questions
about whether it represents a distinct phenomenon or simply a natural
extension of team science. Therefore, we originally arbitrarily chose a
definition for BTS publication which we had decided to be publications
with at least ten authors at ten different institutions that were
published in peer-reviewed journals or had posted a full paper
pre-print. While this definition is a somewhat arbitrary choice, we
separate this research from research on team science that uses any
multi-university collaboration as a definition [@jones2008] to focus on
larger sized teams rather than teams of any size. With at least ten
institutions, the complexities of infrastructure, resources, tenure and
promotion policies, ethics review, and more can occur [@forscher2022a].
Therefore, we believe this choice selects publications that would be
"big" teams and those potential obstacles.

This manuscript seeks not only to evaluate whether this definition is an
appropriate fit for BTS but also to explore whether the criteria for
defining BTS should vary by subject area or evolve over time.
Understanding these nuances will help clarify whether BTS represents a
fundamentally new approach to collaboration or a natural progression of
team science in response to global and technological trends.

We then will examine if there is a true difference between “Team
Science” and "Big Team Science”. Therefore, the goal of this manuscript
is to examine both the *publications* and *people* involved in BTS
projects and “Team Science.” We present descriptive information about
the publication sources and types of articles to demonstrate large-scale
research. Next, we examine the individuals involved in BTS and “Team
Science” for descriptive and predictive purposes. To describe the people
involved in BTS projects vs “Team Science”, we planned to use types of
publications from individuals. For predictive statistics, we explored
the change in diversity of authors over time. It is unclear if the focus
of de-WEIRDing science has only focused on the representation of the
research participants or if it has also improved the representation of
researchers outside of North America and Europe. Last, we examined for a
change in diversity within first author(s) and the last author across
time. As hiring and promoting practices often place a heavy weight on
publications and especially “influential” publications, it becomes
necessary to critically examine the representation present in authorship
in BTS and Team Science” projects.

@@

```{r stats-authors-fig, include = F, fig.cap="The left panel depicts the number of authors included on a paper by subject area, and the right panel demonstrates the number of affiliations by subject area. The boxplot shows the median (bold line), the interquartile range (the box), and the minimum to the 90th percentile of the number of authors/affiliations as the range line. Normally these plots include the entire range of the data, but these extreme range made the boxplot information unreadable. The dots indicate the average number of authors/affiliations for each area with the size of the dot indicating the standard deviation of the statistic. Therefore, larger dots indicate more variability in the number of authors and affiliations. "}
temp_graph <- DF$author_stats %>% 
  filter(SubjectArea != "Multidisciplinary")

@@

\@ref(fig:stats-authors-fig) displays the number of authors and affiliations by subject area. The figure demonstrates that the median number of authors is largest for health sciences, followed by life science, physical sciences, and then social sciences. The general pattern of team authorship includes about 2-8 authors, from about 2-4 institutions. We used the maximum (i.e., across all subject areas) 90th percentile as our exploratory definition for big team papers after examining the results from this analysis. We selected this percentile to have the high of the distribution, but also to be able to include enough papers for analysis across time. Therefore, big teams were defined as 11 authors from at least 6 different institutions[^In a previous version of this manuscript, we defined big teams as 10 authors from at least 10 institutions based on our own experiences working with research consortium. All definitions are likely subjective, but the definition in this manuscript represents the top 10% of author and affiliations in a large body of papers].

NEED TO RUN ALL TEAMS SUBJECT AREA GRAPH

@@

\@ref(tab:big-teams-table) includes the number of distinct authors and papers for each subject area by overall teams and big teams using our 90th percentile definition. The total number of distinct authors for big team papers was `r big_teams_papers` with `r big_teams_authors` distinct authors. In RQ2, we split the big team data into small (6-10), medium, and large big team grouping for convenience to display/analyze geopolitical regions. The table shows the number of authors and papers for those analyses.

@@

For analyzing changes across time, we split the data into regular teams (2-10 authors, 2-5 affiliations) and big teams (as defined above, 11+ authors, 6+ affiliations). The number of papers found in Scopus across time for each subject area are displayed in Figure \@ref(paper-time-fig). The visual results indicated that the number of regular team papers was increasing the most in physical sciences for all manuscripts, followed by life and health sciences, and the last is social sciences. Examining only big teams shows that the rate is also increasing across time. All teams appear to start increasing in the 1990s, while big teams do not start increasing off floor effects until past 2000. The health and life sciences show the largest increases across time in big teams with the smallest trend in the social sciences.

@@

p1 <- ggplot(temp %>% 
         filter(group == "regular"), 
         aes(x = Year, y = Number_Papers, 
             colour = SubjectArea)) + 
  geom_point() + 
  theme_minimal() + 
  scale_color_discrete(name = "Subject Area") + 
  theme(legend.position = "bottom",
        axis.text.y = element_text(angle = 45, hjust = 1)) + # Angles the y-axis text
  ylab("Number of Papers") + 
  ggtitle("Regular Teams")

p2 <- ggplot(temp %>% 
         filter(group == "big"), 
         aes(x = Year, y = Number_Papers, 
             colour = SubjectArea)) + 
  geom_point() + 
  theme_minimal() + 
  scale_color_discrete(name = "Subject Area") + 
  theme(legend.position = "bottom",
        axis.text.y = element_text(angle = 45, hjust = 1)) + # Angles the y-axis text
  ylab(NULL) +  # Removes the y-axis label
  ggtitle("Big Teams")

@@

ggplot(corr.DF, aes(x = SubjectArea, y = estimate, color = group)) + 
  geom_point() +
  theme_minimal() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = .2) + 
  ylab("Rate Estimate") + 
  xlab("Subject Area") + 
  scale_color_discrete(name = "Team Size", 
                       labels = c("Big", "Regular")) + 
  theme(legend.position = "bottom") + 
  geom_hline(yintercept = 0)
```

Using the `minpack.lm` library [CITE ME], we calculated the exponential rate of growth for regular teams and big teams, and these results are shown in \@ref(fig:rate-estimates-fig). All growth rate confidence intervals excluded zero, indicating an exponential increase in the number of team papers over time. Big team growth rates were always higher than their regular team counterparts, but the 95% confidence intervals for the growth estimate overlapped for all statistics. Therefore, the growth trends, while visually appearing to be different, were likely similar for each subject area and team size when examined by estimating exponential growth statistics.

@@

Figure \@ref(fig:fig-career) portrays the average career length for
authors involved in BTS publications across years. Career length was
defined as the year of first publication minus the current year, and
higher numbers mean longer careers. To analyze trends over time, we
calculated the average career length for each publication (i.e., averaging author career lengths to create one score for each paper) and analyzed a
regression analysis using career length to predict year of publication.
In order to show variance between individuals, we calculated the
standard deviation of career length for each publication and used this
variance as an additional predictor.

Negative career length slopes would indicate more young scholars in
later years (i.e., lower average career length as time increases).
Positive career length slopes would indicate older scholars in later
years (i.e., higher average career length as time increases). Negative
career variance slopes imply that variability decreases over the years,
so the average career length is more homogeneous. Positive career length
slopes imply that variability increases over the years, so the average
career length is varied across individuals (i.e., different stages of
scholars). Figure \@ref(fig:fig-heatmap) displays the results for all
regression analyses to compare coefficient strength across and within
each hypothesis.

### Edit me

@@

```{r fig-career, include = TRUE,  warning = FALSE, message = FALSE, fig.cap = "Average career length for big-team science authors. Larger dots indicate more variability in career length for authors by averaging the standard deviation in career length for each manuscript within a year. The data has been filtered to at least 10 publications in a year for this graph.", fig.width=10, fig.height=6}
ggplot(DF$career_time %>% 
         filter(SubjectArea != "Multidisciplinary") %>% 
         filter(num_papers >= 10) %>% 
         filter(Year < 2024) %>% 
         mutate(Team = factor(Team, 
                              levels = c("Regular", "Big"))), 
       aes(Year, avg_career, color = Team)) +
  theme_classic(base_size = 15) + 
  geom_point(aes(size = sd_career)) +
  # geom_point() + 
  facet_wrap(~SubjectArea) + 
  theme(legend.position = "bottom") + 
  ylab("Average Career Length") +
  xlab("Year of Publication") + 
  scale_color_brewer(palette = "Dark2", name = "Subject Area") +
  guides(size = "none")  
  #scale_size_continuous(name = "Number Papers")

```{r career-results, include = FALSE}

label <- c("Physical", "Social", "Health", "Life")
avg_b_values <- c(-0.692014, -0.365357, -0.451827, -0.508350)
avg_se_values <- c(0.002727, 0.005525, 0.001348, 0.001786)
sd_b_values <- c(-0.265538, -0.097108, -0.096730, -0.013958)
sd_se_values <- c(0.007018, 0.013409, 0.003451, 0.004523)
r2_values <- c(0.465446, 0.254312, 0.303781, 0.350923)
sample_size <- total_num

@@

graph_data$analysis <- "Career"

@@

```{r fig-heatmap, include = TRUE, fig.cap="Heatmap results of regression analyses for career length, number of publications, and geopolitical diversity within the region. Each square represents a *b* value or the slope of the predictor (x-axis) onto the dependent variable (each panel), with the exception of the bottom row which is the effect size of each regression analysis $R^2$. Slopes included both the overall value of the predictor ($b$, $b_M$) and the standard deviation of the predictor over time ($b_{SD}$). The color of the square represents the strength of the predictor. The top figure represents all results together for comparison across analyses. The bottom row represents individual heatmaps for each hypothesis to distinguish small differences between subject areas for those research questions. Non-significant results are indicated with NS on the plot.", warning = FALSE, message = FALSE, fig.width=10, fig.height=6}
graph_overall <- graph_data %>% 
  bind_rows(graph_data_2) %>% 
  bind_rows(graph_data_3) %>% 
  mutate(type = factor(type, levels = c("Effect Size", "SD Coefficient", "Coefficient"))) %>% 
  mutate(Estimate = estimate) %>% 
  mutate(heat_label = ifelse(
    low < .00001 & high > .00001, "NS", ""
  ))

>>>>

header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip @plus 0.2ex @minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{@startsection{subparagraph}{5}{1em}%
      {0\baselineskip @plus 0.2ex @minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

@@

# things to import
var_list <- c("author_stats", "paper_stats", "author_paper_year",
              "career_time", "lm_coefficients")

@@

Scientific discovery has increasingly become a collaborative process,
and the scale and scope of team science have dramatically expanded in
recent years [@council2015]. Collaboration in scientific endeavors
involves multiple researchers at (potentially) multiple institutions to
communicate and work together to advance knowledge in their chosen
field(s). Collaboration can manifest uniquely in each project dependent
on the skill sets, hypotheses, and perspectives of collaborators. One
key aspect of collaboration is the flexibility allotted as it is shaped
by the needs of the project and the researcher involved. While
collaboration is not new in science, interest in "team science" is
growing as individual researchers seek an interdisciplinary approach to
research or bring on more students to their project. Team science is
often defined as groups of researchers with various expertise working
together to investigate complex problems [@fiore2008]. Unlike general
collaboration, team science involves structured roles, coordinated
workflows, and shared resources to address challenges that would be
difficult for individual researchers or one small team to solve
independently.

The movement toward team science reflects demands of modern research to
answer increasingly complex questions, meet funder and university desire
for interdisciplinary research, and the desire to increase diversity of
scientific representation [@council2015]. Further, the evolution of team
science reflects broader shifts in research practices, driven by two
sources: 1) increasing globalization and technology that allows for
real-time interdisciplinary research [@jones2008], and 2) expanding
interest in reproducibility, replication, and generalizability [i.e.,
the credibility movement, @maxwell2015; @nelson2018; @zwaan2018;
@vazire2022]. Technological advances have provided easier ways to
collaborate with people who are from other universities and countries
through document sharing platforms (e.g., Google, GitHub, and the Open
Science Framework), video chatting platforms (e.g., Zoom, Microsoft
Teams), and messaging and project management platforms (e.g., Slack,
Trello, when2meet, etc.).

The credibility movement has been the main catalyst for the increased
interest and broader shift of research practices. The movement
emphasizes reproducibility and transparency in science, encouraging
researchers to form new ways to increase the rigor in scientific
endeavors. As the movement has adapted throughout the last decade, the
idea to form larger, more diverse teams and to include participants from
varied backgrounds has become even more a goal to not only increase the
credibility but the generalizability and reliability of scientific
findings. This form of collaboration has been coined “Big Team Science.”
Big Team Science (BTS) builds on team science by scaling efforts to
include larger, often globally diverse teams, which requires significant
coordination and infrastructure [@coles2022; @forscher2022a;
@stewart2017]. BTS projects and organizations organize extensive
collaborations, intentionally incorporating diverse populations and
perspectives into research. This large-scale approach enhances the
reliability and generalizability of findings by integrating varied
methodologies and viewpoints, leading to more robust and inclusive
scientific outcomes. BTS organizations often pool extensive networks of
researchers and resources, aiming to tackle grand scientific challenges
that would be difficult to address within smaller or less coordinated
collaborations. By having both collaborations that span across the globe
and subfields of research areas, age groups, and education levels should
help to drive science in the path of better materials, reliability,
generalizability, and more robust sample sizes (when necessary) in a
study [@auspurg2021; @nosek2014method; @lebel2018].

For example, psychology has had an increase in BTS publications like the
Open Science Collaboration [@opensciencecollaboration2015], Many Labs
Collaborations [@buttrick2020; @ebersole2016; @ebersole2020; @klein2018;
@klein2022; @mathur2020; @skorb2020] or the first papers from the
Psychological Science Accelerator [@bago2022; @buchanan2023;
@dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@moshontz2018; @wang2021]. The success and interest in the large-scale
reproducibility projects [@opensciencecollaboration2015;
@10.7554/eLife.71601], paired with the meta-scientific publications
focusing on researcher practices and incentive structures
[@silberzahn2018many; @john2012] led to a change in journal guidelines
and incentives for researchers interested in participating in
large-scale studies overall [@grahe2014; @kidwell2016; @nosek2015;
@mayo-wilson2021]. In some fields, the BTS movement demonstrated that
large-scale teams were a practical (and publishable) solution to
answering research questions in generalizable way. The support for
Registered Reports, papers accepted before the data has been collected
[@nosek2014; @stewart2020], has allowed researchers to invest in
projects that they know should be published when the project is
complete. Further, the implementation of the Transparency and Openness
Guidelines [@nosek2015] and the Contributor Role Taxonomy (CRediT)
system [@allen2019] have pushed journals and researchers to promote more
open, inclusive publication practices.

Beyond replication concerns, the credibility movement has mirrored calls
for diversification or de-WEIRDing (e.g., Western, Educated,
Industrialized, Rich, and Democratic) scientific research [@henrich2010;
@newson2021; @rad2018] by improving representation in research samples.
Like the large-scale studies in Physics [@aphilos2021; @castelnovo2018]
and Biology [@collins2003], the Social Sciences struggle to represent
the breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator [@moshontz2018], ManyBabies
(<https://manybabies.github.io/>), NutNet (<https://nutnet.org/>), and
DRAGNet (<https://dragnetglobal.weebly.com/>) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more [@bago2022;
@dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@vanbavel2022; @wang2021]. While these organizations and one-time groups
for BTS studies have provided an incredible wealth of data for the
scientific community, we do not yet know exactly how to define BTS -
which is generally termed "an unusually large number of collaborators"
[@forscher2022a; @coles2022].

The lack of formal definition raises questions about whether it
represents a distinct phenomenon or simply a natural extension of team
science. These big teams pose unique challenges, including coordinating
work across diverse time zones, managing conflicts in decision-making,
and ensuring fair distribution of credit for contributions
[@cummings2007; @wuchty2007], but also could provide big rewards by
pooling expertise and increased interdisciplinary funding [@fiore2008].
This paper seeks to clarify the concept of BTS by first establishing a
data-driven definition based on publication patterns. With this
quantitative distinction in place, we analyze publication trends over
time to assess the trajectories of both traditional team science and
BTS. Furthermore, we investigate the diversity of authors involved in
these collaborations to explore whether shifts in the scientific
landscape—such as efforts to de-WEIRD science and the expansion of
collaborative opportunities—have influenced who participates in team
science and BTS. By synthesizing insights from the growth and
diversification of team science, this paper seeks to critically examine
the emergence of big teams. Specifically, it aims to explore whether big
teams are quantitatively different from traditional collaboration models
with the following research questions.

@@

```{r stats-authors-fig, include = F, fig.cap="The left panel depicts the number of authors included on a paper by subject area, and the right panel demonstrates the number of affiliations by subject area. The boxplot shows the median (bold line), the interquartile range (the box), and the minimum to the 90th percentile of the number of authors/affiliations as the range line. Normally these plots include the entire range of the data, but these extreme range made the boxplot information unreadable. The dots indicate the average number of authors/affiliations for each area with the size of the dot indicating the standard deviation of the statistic. Therefore, larger dots indicate more variability in the number of authors and affiliations."}
temp_graph <- DF$author_stats %>% 
  filter(SubjectArea != "Multidisciplinary")

@@

\@ref(fig:stats-authors-fig) displays the number of authors and
affiliations by subject area. The figure demonstrates that the median
number of authors is largest for health sciences, followed by life
science, physical sciences, and then social sciences. The general
pattern of team authorship includes about 2-8 authors, from about 2-4
institutions. We used the maximum (i.e., across all subject areas) 90th
percentile as our exploratory definition for big team papers after
examining the results from this analysis. We selected this percentile to
have the high of the distribution, but also to be able to include enough
papers for analysis across time. Therefore, big teams were defined as 11
authors from at least 6 different institutions[\^In a previous version
of this manuscript, we defined big teams as 10 authors from at least 10
institutions based on our own experiences working with research
consortium. All definitions are likely subjective, but the definition in
this manuscript represents the top 10% of author and affiliations in a
large body of papers].

NEED TO RUN ALL TEAMS SUBJECT AREA

@@

\@ref(tab:big-teams-table) includes the number of distinct authors and
papers for each subject area by overall teams and big teams using our
90th percentile definition. The total number of distinct authors for big
team papers was `r big_teams_papers` with `r big_teams_authors` distinct
authors. In RQ2, we split the big team data into small (11-49), medium
(50-99), and large big team (100+) groupings for convenience to
display/analyze geopolitical regions. The table shows the number of
authors and papers for those analyses.

@@

For analyzing changes across time, we split the data into team science
projects (2-10 authors, 2-5 affiliations) and BTS projects (as defined
above, 11+ authors, 6+ affiliations). The number of papers found in
Scopus across time for each subject area are displayed in Figure
\@ref(paper-time-fig). The visual results indicated that the number of
team science papers was increasing the most in physical sciences for all
manuscripts, followed by life and health sciences, and the last is
social sciences. Examining only BTS projects shows that the rate is also
increasing across time. All teams appear to start increasing in the
1990s, while BTS projects do not start increasing off floor effects
until past 2000. The health and life sciences show the largest increases
across time in big teams with the smallest trend in the social sciences.

@@

p1 <- ggplot(temp %>% 
         filter(group == "regular"), 
         aes(x = Year, y = Number_Papers, 
             colour = SubjectArea)) + 
  geom_point() + 
  theme_minimal() + 
  scale_color_discrete(name = "Subject Area") + 
  theme(legend.position = "bottom",
        axis.text.y = element_text(angle = 45, hjust = 1)) + # Angles the y-axis text
  ylab("Number of Papers") + 
  ggtitle("Team Science Projects")

p2 <- ggplot(temp %>% 
         filter(group == "big"), 
         aes(x = Year, y = Number_Papers, 
             colour = SubjectArea)) + 
  geom_point() + 
  theme_minimal() + 
  scale_color_discrete(name = "Subject Area") + 
  theme(legend.position = "bottom",
        axis.text.y = element_text(angle = 45, hjust = 1)) + # Angles the y-axis text
  ylab(NULL) +  # Removes the y-axis label
  ggtitle("BTS Projects")

@@

ggplot(corr.DF, aes(x = SubjectArea, y = estimate, color = group)) + 
  geom_point() +
  theme_minimal() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = .2) + 
  ylab("Rate Estimate") + 
  xlab("Subject Area") + 
  scale_color_discrete(name = "Team Size", 
                       labels = c("BTS", "Team Science")) + 
  theme(legend.position = "bottom") + 
  geom_hline(yintercept = 0)
```

Using the `minpack.lm` library [@elzhov2023], we calculated the
exponential rate of growth for team science and BTS projects, and these
results are shown in \@ref(fig:rate-estimates-fig). All growth rate
confidence intervals excluded zero, indicating an exponential increase
in the number of team papers over time. BTS growth rates were
always higher than their team science counterparts, but the 95%
confidence intervals for the growth estimate overlapped for all
statistics. Therefore, the growth trends, while visually appearing to be
different, were likely similar for each subject area and team size when
examined by estimating exponential growth statistics.

@@

```{r fig-career, include = TRUE,  warning = FALSE, message = FALSE, fig.cap = "Average career length for big-team science authors. Larger dots indicate more variability in career length for authors by averaging the standard deviation in career length for each manuscript within a year. The data has been filtered to at least 10 publications in a year for this graph.", fig.width=10, fig.height=6}
ggplot(DF$career_time %>% 
         filter(SubjectArea != "Multidisciplinary") %>% 
         filter(num_papers >= 10) %>% 
         filter(Year < 2024) %>% 
         mutate(Team = factor(Team, 
                              levels = c("Regular", "Big"),
                              labels = c("Team Science", "BTS"))), 
       aes(Year, avg_career, color = Team)) +
  theme_classic(base_size = 15) + 
  geom_point(aes(size = sd_career)) +
  facet_wrap(~SubjectArea) + 
  theme(legend.position = "bottom") + 
  ylab("Average Career Length") +
  xlab("Year of Publication") + 
  scale_color_manual(values = c("#CA0020", "#0571B0"), name = "Team Size") + # Red and Blue from RdBu ends
  guides(size = "none")
  #scale_size_continuous(name = "Number Papers")

Figure \@ref(fig:fig-career) portrays the average career length for
authors involved in team science and BTS projects over time. Career length was defined as the year of first publication minus the current year, and higher numbers mean longer careers. The general pattern for team science and BTS projects is a decrease in average career length over time. However, it appears that, in at least the last two decades, BTS projects average a longer career length than team science projects. This trend is visually consistent across all four subject areas examined.

To analyze these trends over time, we calculated the average career length for each publication (i.e., averaging author career lengths to create one score for each paper) and analyzed a regression analysis using career length to predict year of
publication. In order to show variance between individuals, we
calculated the standard deviation of career length for each publication
and used this variance as an additional predictor. Negative career length slopes would indicate more young scholars in
later years (i.e., lower average career length as time increases).
Positive career length slopes would indicate older scholars in later
years (i.e., higher average career length as time increases). Negative
career variance slopes imply that variability decreases over the years,
so the average career length is more homogeneous. Positive career length
slopes imply that variability increases over the years, so the average
career length is varied across individuals (i.e., different stages of
scholars). Figure \@ref(fig:fig-heatmap) displays the results for all
regression analyses to compare coefficient strength across and within
each hypothesis.

```{r lm-coefficients}
sample_size <- 1000000
DF$lm_coefficients$r2_low <- DF$lm_coefficients$r2_high <- NA

# add upper and lower Rsquared values
for (i in 1:nrow(DF$lm_coefficients)){
  temp <- ci.R2(R2 = DF$lm_coefficients$rsquared[i], 1, sample_size)
  DF$lm_coefficients$r2_low[i] <- temp$Lower.Conf.Limit.R2
  DF$lm_coefficients$r2_high[i] <- temp$Upper.Conf.Limit.R2
}

# create upper and lower CIs for each coefficient
DF$lm_coefficients <- DF$lm_coefficients %>% 
  mutate(b_avg_low = b_avg_effect - qt(.025, sample_size, lower.tail = F)*se_avg_effect,
         b_avg_high = b_avg_effect + qt(.025, sample_size, lower.tail = F)*se_avg_effect,
         b_std_low = b_std_effect - qt(.025, sample_size, lower.tail = F)*se_std_effect,
         b_std_high = b_std_effect - qt(.025, sample_size, lower.tail = F)*se_std_effect)

```

### Edit me starting here ###

@@

```{r career-results-teams, include = FALSE}
label <- c("Physical", "Social", "Health", "Life")
avg_b_values <- c(-0.692014, -0.365357, -0.451827, -0.508350)
avg_se_values <- c(0.002727, 0.005525, 0.001348, 0.001786)
sd_b_values <- c(-0.265538, -0.097108, -0.096730, -0.013958)
sd_se_values <- c(0.007018, 0.013409, 0.003451, 0.004523)
r2_values <- c(0.465446, 0.254312, 0.303781, 0.350923)
sample_size <- total_num <- 10000000

@@

graph_data$analysis <- "Career"
```

```{r career-results-bts, include = FALSE}
label <- c("Physical", "Social", "Health", "Life")
avg_b_values <- c(-0.692014, -0.365357, -0.451827, -0.508350)
avg_se_values <- c(0.002727, 0.005525, 0.001348, 0.001786)
sd_b_values <- c(-0.265538, -0.097108, -0.096730, -0.013958)
sd_se_values <- c(0.007018, 0.013409, 0.003451, 0.004523)
r2_values <- c(0.465446, 0.254312, 0.303781, 0.350923)
sample_size <- total_num

graph_data_bts <- data.frame(
  "label" = label,
  "estimate" = avg_b_values,
  "type" = "Coefficient",
  "low" = avg_b_values - qt(.025, sample_size, lower.tail = F)*avg_se_values,
  "high" = avg_b_values + qt(.025, sample_size, lower.tail = F)*avg_se_values) %>% 
    bind_rows(
    data.frame(
    "label" = label,
    "estimate" = sd_b_values,
    "type" = "SD Coefficient",
    "low" = sd_b_values - qt(.025, sample_size, lower.tail = F)*sd_se_values,
    "high" = sd_b_values + qt(.025, sample_size, lower.tail = F)*sd_se_values) 
    ) %>% 
  bind_rows(
    data.frame(  
      "label" = label, 
      "estimate" = r2_values,
      "type" = "Effect Size",
      "low" = r2_low,
      "high" = r2_high)
  )

graph_data_bts$type <- factor(
  graph_data_bts$type, 
  levels = c("Coefficient", "SD Coefficient", "Effect Size")
)

graph_data_bts$analysis <- "Career"
```

@@

```{r fig-heatmap}

, include = TRUE, fig.cap="Heatmap results of regression analyses for career length, number of publications, and geopolitical 
within the region. Each square represents a *b* value or the slope of the predictor (x-axis) onto the dependent variable (each panel), with the exception of the bottom row which is the effect size of each regression analysis $R^2$. Slopes included both the overall value of the predictor ($b$, $b_M$) and the standard deviation of the predictor over time ($b_{SD}$). The color of the square represents the strength of the predictor. The top figure represents all results together for comparison across analyses. The bottom row represents individual heatmaps for each hypothesis to distinguish small differences between subject areas for those research questions. Non-significant results are indicated with NS on the plot.", warning = FALSE, message = FALSE, fig.width=10, fig.height=6

graph_overall <- graph_data %>% 
  bind_rows(graph_data_2) %>% 
  bind_rows(graph_data_3) %>% 
  mutate(type = factor(type, levels = c("Effect Size", "SD Coefficient", "Coefficient"))) %>% 
  mutate(Estimate = estimate) %>% 
  mutate(heat_label = ifelse(
    low < .00001 & high > .00001, "NS", ""
  ))
